---
title: "Georgia Election Map"
author: "Kristle Chester"
date: "November 22, 2020"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(maps)
library(tidyverse)
library(readxl)
library(ggplot2)
library(plyr)
library(dplyr)
library(sf)
library(spatstat)
library(maptools) #Needed for as.owin
library(kableExtra)
library(scales)
#library(purrr)
#library(gganimate)
#library(gifski)
library(xml2)
library(extrafont)



# Load workspace with fonts
load("fonts.Rdata")


## setting working directory
knitr::opts_knit$set(root.dir = "C:/Users/kcyar/OneDrive - The Index Writer, LLC/Medium/better_ga_map")

```

## Voter Density Matters

```{r xml_load, echo=FALSE}

load("turnout.rData")
 
```

## Map

```{r map, echo=FALSE}
load("ga_map.Rdata")
load("ga_clean.Rdata")


```

```{r voter_pts, echo=FALSE, results = FALSE, message=FALSE, warning=FALSE}

 # Generate points within precinct to represent voter density. Note, this uses Ballots_Cast, not Total_Votes due to the previous issues mentioned with Hall County. In truth,100% voter turnout is practically unheard. On election day, the results are all about who shows up. Who stays home makes a difference, but we're visualizing who showed up because that is the election result.

# Generating a random point for each vote within a precinct takes about 1/2 an hour. There is no need to repeat this step everytime you knit the rmd file.

# If the xml data changes, rerun create_voter_points.R.
load("voter_points.Rdata")

```


```{r density, echo=FALSE}

# All election_merge ga_map. Note, 2,656 precincts reported results for the 2020 general election. 

all_merged <- ga_map %>%
  st_as_sf()


# R_MOV_BC = 100 is Red. R_MOV_BC = -100 is BLUE

# Choropleth Maps
this_rb <- c("blue4", "blue4", "blue3", "blue2", "#80a0ff", "#ff8080", "red2", "red3", "red4", "red4")

# Alternate colors.

#this_colorblind <- c("#431A53", "#572F67", "#6E477D", "#896996", "#AD97B6", "#76DB76", "#4CC94C", "#2CB72C", "#129C12", "#037C03")

this_colorblind <- c("#037C03", "#129C12", "#2CB72C", "#4CC94C", "#76DB76", "#AD97B6", "#896996", "#6E477D", "#572F67", "#431A53")



this_rescale <- c(
  0,
  0.175, # D 65 
  0.33875, # D 32.25
  0.419375, # D 16.125
  0.4975, # D 0.5
  0.5025, # R 0.5
  0.580625, # R 16.125
  0.66125, # R 32.25
  0.825, # R 65
  1
)
#this_rescale <- c(-70, -50, -25, -5, -3, 0, 3, 5, 25, 50, 70)
this_break <- c(
 # 0.1, # blue4
  0.175, # D 65
  0.3, # D 40
  0.425, # D 15
  0.575,# #R 15
  0.7, # R 40
  0.825#, # r 65
  #0.9# red4
  )
party_labels <- c("D +65", "D +40", "D +15", "R +15", "R +40", "R +65")
this_no_party <- c("B +65", "B +40", "B +15", "A +15", "A +40", "A +65")

choro_map <- function(this_color, map_colors, this_labels, this_title, this_subtitle) {
  
  this_color_test <- deparse(substitute(this_color))

  if(this_color_test == "all_merged") {
    outlines = NA
  } else {
    outlines = "gray30"
  }

  ga_color <- ggplot()

  ga_color + geom_sf(data=ga_clean, fill="gray30", color="gray30", size=1.1) +
    geom_sf(data=this_color, aes(fill=R_MOV_RDL_COLOR),  color=outlines, size=0.5, stroke = 0, shape=20) +
    scale_fill_gradientn(colors=map_colors, breaks=this_break, labels=this_labels, values = scales::rescale(this_rescale), limits = c(0.0, 1.0)) +
    labs(title=this_title, subtitle=this_subtitle) +
    theme_void() +
    theme(legend.title =  element_blank(),
          legend.text = element_text(size=18, color="gray30", family="Noto Sans"),
          plot.title = element_text(hjust = 0.5, size = 35, family="Noto Sans"),
          plot.subtitle = element_text(hjust = 0.5, size = 30, color="gray30", vjust = -1, face="italic", family="Noto Sans")
          ) +
    guides(fill = guide_colourbar(barwidth=2, barheight = 20, ticks.linewidth = 3.5))
}

##  Precinct Map.

ga_pre_map <- choro_map(all_merged, this_rb, party_labels, "Georgia General Election 2020", "Presidential by Precinct")
ggsave("output/ga_precinct_choropleth.png", ga_pre_map, width = 14, height=14, units="in")

## County Map
ga_all_county_map <- choro_map(ga_county_map, this_rb, party_labels, "Georgia General Election 2020", "Presidential by County")
ggsave("output/ga_county_choropleth.png", ga_all_county_map, width = 14, height=14, units="in")

# Colorblind Palette
##  Precinct Map.

ga_pre_map_cb <- choro_map(all_merged, this_colorblind, this_no_party, "Some Random Election in Georgia", "By Precinct")
ggsave("output/ga_precinct_choropleth_cb.png", ga_pre_map_cb, width = 14, height=14, units="in")

## County Map
ga_all_county_map_cb <- choro_map(ga_county_map, this_colorblind, this_no_party, "Some Random Election in Georgia", "By County")
ggsave("output/ga_county_choropleth_cb.png", ga_all_county_map_cb, width = 14, height=14, units="in")


```

```{r density_mapping, echo=FALSE}

voter_dot_map <- function(scale, map_colors, this_labels, this_title, this_subtitle) {
  
  
  # Extract x y coordinates
  voter_dots_coord <- st_coordinates(scale)
# Bind the data back together

  voter_dots_xy <- cbind(scale, voter_dots_coord)

# Vote Density at 100 voters per dot.

  ga_density <- ggplot()

  ga_density + geom_sf(data=ga_clean, fill="white", color="gray30", size=1.1) +
    geom_point(data=voter_dots_xy, aes(X, Y, color=R_MOV_RDL_COLOR), size=0.5, stroke = 0, shape=20) +
    scale_color_gradientn(colors=map_colors, breaks=this_break, labels=this_labels, scales::rescale(this_rescale), limits = c(0.0, 1.0)) +
    labs(title=this_title, subtitle=this_subtitle, color="Margin of Victory") +
    theme_void() +
    theme(legend.title =  element_blank(),
          legend.text = element_text(size=18, color="gray30", family="Noto Sans"),
          plot.title = element_text(hjust = 0.5, size = 35, family="Noto Sans"),
          plot.subtitle = element_text(hjust = 0.5, size = 30, color="gray30", vjust = -1, face="italic", family="Noto Sans")
          ) +
    guides(color = guide_colourbar(barwidth=2, barheight = 20, ticks.linewidth = 3.5))
}

#guide = guide_legend(keyheight=2)

#ga_map_voter_100 <- voter_dot_map(voter_dots_100)

#ggsave("output/voter_density_100.png", ga_map_voter_100)

#ga_map_voter_50 <- voter_dot_map(voter_dots_50)
#ggsave("output/voter_density_50.png", ga_map_voter_50)

#ga_map_voter_25 <- voter_dot_map(voter_dots_25)
#ggsave("output/voter_density_25.png" ,ga_map_voter_25)

ga_map_voter_10 <- voter_dot_map(voter_dots_10, this_rb, party_labels, "Georgia General Election 2020", "Presidential")

ggsave("output/voter_density_10.png", ga_map_voter_10, width = 14, height=14, units="in")

#ga_map_voter_1 <- voter_dot_map(voter_dots_1)
#ggsave("output/voter_density_all.png", ga_map_voter_1, width = 46, height=46, units="in")

# Colorblind
ga_map_voter_10_cb <- voter_dot_map(voter_dots_10, this_colorblind, this_no_party, "Some Random Election in Georgia", "")

ggsave("output/voter_density_10_cb.png", ga_map_voter_10_cb, width = 14, height=14, units="in")

```

```{r add_stats, echo=FALSE}

percentile_90 <- election_merge %>%
  filter(Total_Voters >= quantile(Total_Voters, 0.9))
total_voters_90 <- format(sum(percentile_90$Total_Voters), scientific=FALSE)


percentile_95 <- election_merge %>%
  filter(Total_Voters >= quantile(Total_Voters, 0.95))
total_voters_95 <- format(sum(percentile_95$Total_Voters), scientific=FALSE)


percentile_98 <- election_merge %>%
  filter(Total_Voters >= quantile(Total_Voters, 0.98))
total_voters_98 <- format(sum(percentile_98$Total_Voters), scientific=FALSE)


percentile_99 <- election_merge %>%
  filter(Total_Voters >= quantile(Total_Voters, 0.99))
total_voters_99 <- format(sum(percentile_99$Total_Voters), scientific=FALSE)


r_election_merge <- election_merge %>%
  filter(R_Percentage_BC > D_Percentage_BC)

r_undervote <- sum(r_election_merge$Undervotes)

d_election_merge <- election_merge %>%
  filter(D_Percentage_BC > R_Percentage_BC)

d_undervote<- sum(d_election_merge$Undervotes)

diff_r_d_undervote <- r_undervote - d_undervote 

neutral_election_merge <- election_merge %>%
  filter(D_Percentage_BC == R_Percentage_BC)

calc_undervote_sub <- function(party, x, y) {
  
  if(party == "d") {
    this_election_filter <- election_merge %>%
    filter(x < D_Percentage_BC & D_Percentage_BC < y)
    total_undervote <- sum(this_election_filter$Undervotes)

  } else {
    this_party_df <- r_election_merge
    this_percent <- "R_Percentage_BC"
  }
  
  
}

d70_under <- calc_undervote_sub("d", 0.70, 1.00)

d65_under <- calc_undervote_sub("d", 0.65, 0.70)

d60_under <- calc_undervote_sub("d", 0.6, 0.65)

d55_under <- calc_undervote_sub("d", 0.55, 0.6)

d50_under <- calc_undervote_sub("d", 0.5, 0.55)

r70_under <- calc_undervote_sub("r", 0.70, 1.00)

r65_under <- calc_undervote_sub("r", 0.65, 0.70)

r60_under <- calc_undervote_sub("r", 0.6, 0.65)

r55_under <- calc_undervote_sub("r", 0.55, 0.6)

r50_under <- calc_undervote_sub("r", 0.5, 0.55)



#r75 <- election_merge %>%
 # filter(R_MOV_BC >= 75)

total_registered_voters <- sum(election_merge$Total_Voters)

total_election_votes <- sum(election_merge$Ballots_Cast)

precincts_reporting <- as.numeric(nrow(election_merge))

election_overview <- data.frame("Total_Registered_Voters" = total_registered_voters, "Total_Ballots_Cast" = total_election_votes, "Voter_Turnout" = round(total_election_votes/total_registered_voters, 4), "Precincts Total" = precincts_reporting)

tallies <- function(this_race, this_election_df) {
  
  this_total_votes <- sum(this_election_df[,12]) + sum(this_election_df[,17]) + sum(this_election_df[,22])
  
  this_total_undervote <- sum(this_election_df$Undervotes)
  
  this_total_overvote <- sum(this_election_df$Overvotes)
  
  this_rep_margin <- sum(this_election_df[,12]) - sum(this_election_df[,17])
  
  this_tallies <- data.frame("Race" = this_race, "Total_Undervotes" = this_total_undervote, "Percent_Undervotes" = round(this_total_undervote/total_election_votes, 4), "Total_Overvote" = this_total_overvote, "Republican" = sum(this_election_df[,12]), "Republican_Percent" = round(sum(this_election_df[,12])/total_election_votes, 4), "Democratic" = sum(this_election_df[,17]), "Democratic_Percent" = round(sum(this_election_df[,17])/total_election_votes, 4), "Libertarian" = sum(this_election_df[,22]), "Libertarian_Percent" = round(sum(this_election_df[,22])/total_election_votes,4), "Republican Margin" = this_rep_margin, "Difference Between Libertarian and the Margin" = sum(this_election_df[,22]) - abs(this_rep_margin) )
}

pres_totals <- tallies("Presidential", election_merge)

# Load the Perdue senate race data.

load(file="sen1_data.Rdata")

senate1_totals <- tallies("Senate - Perdue", senate1_merge)

election_totals <- rbind(pres_totals, senate1_totals)

diff_between_r_pres_sen <- pres_totals[,5] - senate1_totals[,5]

diff_between_d_pres_sen <- pres_totals[,7] - senate1_totals[,7]

diff_between_l_pres_sen <- pres_totals[,9] - senate1_totals[,9]

diff_between_undervotes_pres_sen <- pres_totals[,2] - senate1_totals[,2]

# Write select tables to csv. Some of these are too long to use and will need a little reformatting.

write.csv(election_totals, "output/election_totals.csv")
write.csv(pres_totals, "output/pres_totals.csv")
write.csv(election_overview, "output/election_overview.csv")
```

## Basic Election Stats

`r kable(election_overview)`



## Presidential Tallies

`r kable(pres_totals)`

## Senate Tallies Perdue

`r kable(senate1_totals)`


## Understand Precinct Size

* 90th Percentile Precincts Total Voters: `r total_voters_90`
* 95th Percentile Precincts Total Voters: `r total_voters_95`
* 98th Percentile Precincts Total Voters: `r total_voters_98`
* 99th Percentile Precincts Total Voters: `r total_voters_99`

```{r bar_charts, echo=FALSE}

total_pres_votes <- pres_totals$Republican + pres_totals$Democratic + pres_totals$Libertarian

# Manually enter totals from the presidential recount. (Due to time constraints, pulling the xml is not feasible.) Source: https://results.enr.clarityelections.com/GA/107231/web.264614/#/summary

recount_rep <- 2461854
recount_dem <- 2473633
recount_lib <- 62229

total_recount_votes <- recount_rep + recount_dem + recount_lib

final_chart_row <- data.frame("Category" = c("Registered Voters", "Ballots Cast", "Republican Votes", "Democratic Votes", "Libertarian Votes"), "Total" = c(total_registered_voters, total_election_votes, pres_totals$Republican, pres_totals$Democratic, pres_totals$Libertarian ))

#final_chart_row <- data.frame("Category" = c("Registered Voters", "Ballots Cast", "Republican Votes", "Democratic Votes", "Libertarian Votes"), "Total" = c(total_registered_voters, total_recount_votes, recount_rep, recount_dem, recount_lib ))

final_chart_row <- final_chart_row %>%
  mutate(Category=fct_relevel(Category, "Registered Voters", "Ballots Cast", "Republican Votes", "Democratic Votes", "Libertarian Votes")) %>%
  mutate("Race" = "Presidential", "Percent Registered" = Total/Total[[1]], "Percent Ballots" = Total/Total[[2]])

final_chart_row$`Percent Ballots`[final_chart_row$`Percent Ballots` > 1] <- NA

final_unlikely <- final_chart_row$Total[1] - final_chart_row$Total[2]

final_chart_row$Chart <- c(final_unlikely, 0, final_chart_row$Total[3], final_chart_row$Total[4], final_chart_row$Total[5])

# Create Labels

label_reg_voter <- paste("Registered Nonvoters\n",(percent(round(final_chart_row$`Percent Registered`[1] - final_chart_row$`Percent Registered`[2], 4))), "\n", format(final_chart_row$Chart[1], big.mark=","), sep="")
label_r_voter <- paste("Trump (R)\n", percent(round(final_chart_row$`Percent Registered`[3], 4), accuracy=0.01), "\n", format(final_chart_row$Total[3], big.mark=","), sep="")
label_d_voter <- paste("Biden (D)\n", percent(round(final_chart_row$`Percent Registered`[4], 4), accuracy=0.01), "\n", format(final_chart_row$Total[4], big.mark=","), sep="")
label_l_voter <- paste("Jorgensen (L)\n", percent(round(final_chart_row$`Percent Registered`[5], 4), accuracy=0.01), "\n", format(final_chart_row$Total[5], big.mark=","), sep="")

final_chart_row$Labels <- c(label_reg_voter, 0, label_r_voter, label_d_voter, NA)

final_percent_ballots <- final_chart_row %>%
  drop_na("Percent Ballots")

final_chart_row <- final_chart_row[-c(2),]

stacked_bar <- function(df, colors, this_title, this_l_label) {
  stacked_bar <- ggplot(df, aes(y=Chart, x=Race)) +
  geom_col(aes(fill=Category), width=0.6) +
  geom_bar(position="fill", stat="identity") +
  scale_fill_manual(values=colors) +
  coord_flip() +
  theme_void() +
  geom_text(aes(label = Labels, group=Category), position=position_stack(vjust=0.5), size=5, family="Noto Sans",  fontface="bold", color ="white") +
  annotate("text", x=0.55, y=0, label=this_l_label, family="Noto Sans") +
  labs(title = "Georgia General Election 2020", subtitle=this_title) +
  theme(legend.position="none",
        plot.title = element_text(hjust = 0.5, size = 20, family="Noto Sans"),
        plot.subtitle = element_text(hjust = 0.5, size = 15, color="gray30", vjust = -1, face="italic", family="Noto Sans")
        )
}

final_all_colors <- c("#666666", "red3", "blue3", "green3")

final_all <- stacked_bar(final_chart_row, final_all_colors, "Presidential - Certified Vote", label_l_voter)


ggsave("output/final_all_stacked_bar_ga.png", final_all, width = 14, height=3.1, units="in")

final_pres_row <- final_chart_row
final_pres_row <- final_pres_row[-c(1),]
final_pres_row$Labels <- NULL
final_pres_row$`Percent Registered` <- NULL
final_pres_row$`Percent Ballots` <- NULL


pres_label_r <- paste("Trump (R)\n", percent(round((final_pres_row$`Total`[1]/total_pres_votes), 4), accuracy=0.01), "\n", format(final_pres_row$Total[1], big.mark=","), sep="")
pres_label_d <- paste("Biden (D)\n", percent(round((final_pres_row$`Total`[2]/total_pres_votes), 4), accuracy=0.01), "\n", format(final_pres_row$Total[2], big.mark=","), sep="")
pres_label_l <- paste("Jorgensen (L)\n", percent(round((final_pres_row$`Total`[3]/total_pres_votes), 4), accuracy=0.01), "\n", format(final_pres_row$Total[3], big.mark=","), sep="")


final_pres_row$Labels <- c(pres_label_r, pres_label_d, NA)

final_pres_colors <- c("red3", "blue3", "green3")

final_pres <- stacked_bar(final_pres_row, final_pres_colors, "Presidential - Certified Vote", pres_label_l)

ggsave("output/final_pres_stacked_bar_ga.png", final_pres, width = 14, height=3.1, units="in")

```

